A community-maintained standard library of population genetic simulations

### 2019-12-19 popSimManu.tex
Updated model table.

\documentclass[12pt,halfline,a4paper]{ouparticle}
\usepackage{amsmath,amssymb}
\usepackage[round]{natbib}
\usepackage{graphicx,rotating}
\usepackage{dsfont}
\usepackage{xspace}

@@

\affil[7]{Department of Human Genetics, McGill University}
\affil[8]{Melbourne Integrative Genomics, School of Mathematics and Statistics, University of Melbourne}
\affil[9]{Department of Mathematical Stochastics, University of Freiburg}
\affil[10]{Department of Genome Sciences, University of Washington}
\affil[11]{The Biodesign Institute and The School of Life Sciences, Arizona State University}
\affil[12]{Department of Human Genetics, David Geffen School of Medicine, University of California, Los Angeles}
\affil[13]{Department of Biology, Stanford University}
\affil[14]{Department of Ecology, Evolution, and Environmental Biology, Columbia University}
\affil[15]{Department of Computational Biology, Cornell University}
\affil[16]{Computer Technologies Laboratory, ITMO University}
\affil[17]{International Laboratory for Human Genome Research, National Autonomous University of Mexico}
\affil[18]{Department of Molecular and Cellular Biology, University of Arizona}
\affil[19]{Big Data Institute, Oxford Unviersity}
\affil[$\star$]{Denotes shared first authorship, listed alphabetically}
\affil[$\dagger$]{Denotes shared senior authorship, listed alphabetically}
\affil[$\aleph$]{Denotes corresponding authors, listed alphabetically}
% JK: didn't bother with the \texttt stuff for stdpopsim in the abstract as it's nearly
% always dropped in production anyway.

@@

\begin{table}[t]
\begin{footnotesize}
\input{model_table.tex}
\end{footnotesize}
\caption{\label{tab:catalog}
Details of the initial set of population models across four species.
\textbf{NOTES: Column names are ID, name (or description maybe?),
reference, CPU time, RAM}
}
\end{table}

@@

Here we have described what we view as the first product from the PopSim Consortium:
the \stdpopsim library. We have founded the Consortium with a number of specific goals in mind:
standardization of simulation within the computational genetics community,
increasing reproducibility and ease of use of complex simulation based results,
community-based development and decision making guiding best practices in population genetics,
and eventual benchmarking of inferential methods.

The \stdpopsim library is the first piece of this effort, in that it allows for rigorous
standardization of complex population genetic simulations. Population genetics, as a field,
has yet to coalesce around a set of standards for the crucial task of method
evaluation, which in our field hinges on simulation. In contrast, other fields such as
structural biology \citep{moult1995large} and machine learning \citep{russakovsky2015imagenet} have a long track record
of standardized method testing. We hope that our efforts represent the beginning of what
will prove to be an equally longstanding and invaluable tradition in population genetics.

Individual research groups are often in the position of having to independently
reimplement complex, previously published demographic models.
Additional layers of realism (e.g., recombination maps) makes this especially daunting.
\stdpopsim aims to lower this barrier
by creating a quality controlled, easy-to-use, gateway to models and organisms
that are central to modern research in genetics.

We have illustrated in this paper how \stdpopsim can be used for direct
comparisons of inferential methods on a common set of simulations. Our
benchmarking comparisons have been limited, but nevertheless
reveal some informative comparisons. For
example, at the task of estimating $N(t)$ trajectories for simulations of single human
populations, we find that the sequence-based methods (\MSMC and \smcpp)
perform somewhat better overall---at least for mid-range values of
$t$---than the site frequency spectrum-based method (\stairwayplot)
(Figures~\ref{fig:n_t_ragsdale} and~\ref{fig:n_t_gutenkunst}), which
tends to over-estimate the size of oscillations.  By contrast,
\stairwayplot outperforms the sequence-based methods
on simulations of \textit{D.~melanogaster} populations,
in which linkage disequilibrium is considerably reduced (Figure~\ref{fig:n_t_sheehan}).
In simulations of two human populations
(Figure~\ref{fig:IM_popn_human}), most methods do reasonably well at
reconstructing the YRI history, but struggle with the more complex CEU
history, in large part because of the restriction of constant sizes per
population.  An exception is \smcpp, which does not have the same
restrictions on its inferred history, and as a result does somewhat better
with the CEU history but tends to overfit the YRI history.  The results for
the two-population \textit{D.~melanogaster} case (Figure~\ref{fig:two_popn_fly})
are generally similar.

@@

\section*{Acknowledgements}
We thank the Probabilistic Modeling in Genomics conference organizers with making this collaboration possible.
Early on in the project we were helped by many people including XXXXXX.
CCK and KEL were funded under NIH Award R35GM119856.
JRA and ADK were funded under NIH Award R01GM117241.
FR and GG were supported by a Villum Young Investigator award (project no. 00025300).
DODV is funded by a UC MEXUS-CONACYT Collaborative Grant and a DGAPA-PAPIIT grant (PAPIIT-IA200620).

>>>>

\documentclass[12pt,halfline,a4paper]{ouparticle}
\usepackage{amsmath,amssymb}
\usepackage[round]{natbib}
\usepackage{graphicx,rotating}
\usepackage{dsfont}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{siunitx}

@@

\affil[7]{Department of Human Genetics, McGill University}
\affil[8]{Melbourne Integrative Genomics, School of Mathematics and Statistics, University of Melbourne}
\affil[9]{Department of Mathematical Stochastics, University of Freiburg}
\affil[10]{Department of Genome Sciences, University of Washington}
\affil[11]{The Biodesign Institute and The School of Life Sciences, Arizona State University}
\affil[12]{Department of Human Genetics, David Geffen School of Medicine, University of California, Los Angeles}
\affil[13]{Department of Biology, Stanford University}
\affil[14]{Department of Ecology, Evolution, and Environmental Biology, Columbia University}
\affil[15]{Department of Computational Biology, Cornell University}
\affil[16]{Computer Technologies Laboratory, ITMO University}
\affil[17]{International Laboratory for Human Genome Research, National Autonomous University of Mexico}
\affil[18]{Department of Molecular and Cellular Biology, University of Arizona}
\affil[19]{Big Data Institute, Li Ka Shing Centre for Health Information and
Discovery, University of Oxford}
\affil[$\star$]{Denotes shared first authorship, listed alphabetically}
\affil[$\dagger$]{Denotes shared senior authorship, listed alphabetically}
\affil[$\aleph$]{Denotes corresponding authors, listed alphabetically}
% JK: didn't bother with the \texttt stuff for stdpopsim in the abstract as it's nearly
% always dropped in production anyway.

@@

\renewcommand{\arraystretch}{1.2}
\begin{table}[t]
% Make sure the table is centered even though it's too wide
\makebox[\textwidth][c]{
    \begin{footnotesize}
    \input{model_table.tex}
    \end{footnotesize}
}
\caption{\label{tab:catalog}
Initial set of demographic models in the Catalog and simple benchmarks.
For each model we report the CPU time, maximum memory usage and the
size of the output \tskit file. In each case we simulate 100 samples
drawn from the first population, for the shortest chromosome of that species
and a constant chromosome-specific recombination rate.
The times reported are for a single replicate run on an Intel i5-7600 CPU.
These are not intended to be rigorous benchmarks, and the computing resources
required will vary widely depending on sample sizes, chromosome length,
recombination rates and other factors.
}
\end{table}

@@

Here we have described the first major product from the PopSim Consortium:
the \stdpopsim library. We have founded the Consortium with a number of specific goals in mind:
standardization of simulation within the computational genetics community,
increasing reproducibility and ease of use of complex simulation based results,
community-based development and decision making guiding best practices in population genetics,
and eventual benchmarking of inferential methods.

The \stdpopsim library allows for rigorous
standardization of complex population genetic simulations. Population genetics, as a field,
has yet to coalesce around a set of standards for the crucial task of method
evaluation, which in our field hinges on simulation. In contrast, other fields such as
structural biology \citep{moult1995large} and machine learning \citep{russakovsky2015imagenet} have a long track record
of standardized method testing. We hope that our efforts represent the beginning of what
will prove to be an equally longstanding and valuable tradition in population genetics.

% (ACS) I suggest deleting the following paragraph (commented out below).
% All of the important points are made below (even with some of the same
% phrases, such as "barriers to entry") and I think it is needlessly
% redundant.

%Individual research groups are often in the position of having to independently
%reimplement complex, previously published demographic models.
%Additional layers of realism (e.g., recombination maps) make this especially daunting.
%\stdpopsim aims to lower this barrier
%by creating a quality controlled, easy-to-use, gateway to models and organisms
%that are central to modern research in genetics.

We have illustrated in this paper how \stdpopsim can be used for direct
comparisons of inferential methods on a common set of simulations. Our
benchmarking comparisons have been limited, but nevertheless
reveal some informative features. For
example, at the task of estimating $N(t)$ trajectories for simulated human
populations, we find that the sequence-based methods (\MSMC and \smcpp)
perform somewhat better overall---at least for mid-range values of
$t$---than the site frequency spectrum-based method (\stairwayplot)
(Figures~\ref{fig:n_t_ragsdale} and~\ref{fig:n_t_gutenkunst}), which
tends to over-estimate the sizes of oscillations.  By contrast,
\stairwayplot outperforms the sequence-based methods
on simulations of \textit{D.~melanogaster} populations,
in which linkage disequilibrium is considerably reduced (Figure~\ref{fig:n_t_sheehan}).
In simulations of two human populations
(Figure~\ref{fig:IM_popn_human}), most methods do reasonably well at
reconstructing the YRI history, but struggle with the more complex CEU
history, in large part because of the restriction of constant sizes per
population.  An exception is \smcpp, which does not have the same
restrictions on its inferred history, and as a result does somewhat better
with the CEU history but tends to overfit the YRI history.  The results for
the two-population \textit{D.~melanogaster} case (Figure~\ref{fig:two_popn_fly})
are generally similar.

@@

\section*{Acknowledgements}
We thank the Probabilistic Modeling in Genomics conference organizers with making this collaboration possible.
Early on in the project we were helped by many people including XXXXXX.
CCK and KEL were funded under NIH Award R35GM119856.
JRA and ADK were funded under NIH Award R01GM117241.
TJS and RNG were funded under NIH award R01GM127348.
FR and GG were supported by a Villum Young Investigator award (project no.~00025300).
DODV is funded by a UC MEXUS-CONACYT Collaborative Grant and a DGAPA-PAPIIT grant (PAPIIT-IA200620).
JK is supported by the Robertson Foundation.
